Variational Approximations for Transfer in Reinforcement Learning

### 2018-05-16 NIPS 2018/main.tex
Added MGVT theorem

>>>>

\begin{restatable}{theorem}{thmg} \label{th:main-mixture}
Fix a target task $\tau$ a let $\wt{Q}$ be the fixed-point of the corresponding mellow Bellman operator. Assume linearly parameterized value functions $Q_{\bm{w}}(s,a)=\bm{w}^T\bm{\phi}(s,a)$ with bounded weights $\bm{w}\leq w_{max}$ and uniformly bounded features $\bm{\phi}(s,a)\leq \phi_{\max}$. Consider the mixture version of Alg. \ref{alg:transfer} using $C$ components, source task weights $\mathcal{W}_s$, and prior $p(\bm{w}) =\frac{1}{\abs{\mathcal{W}_s}}\sum_j\mathcal{N}(\bm{w}_j,\sigma^2_p\bm{I})$. Denote by $\wh{\bm{\xi}} = (\wh{\bm{\mu}}_1,\dots,\wh{\bm{\mu}}_C,\wh{\bm{\Sigma}}_1,\dots,\wh{\bm{\Sigma}}_C)$ the variational parameters minimizing the objective of Eq. \ref{eq:elbo} on a dataset $D$ of $N$ samples. Let $\nu$ be a probability measure over $\mathcal{S}\times\mathcal{A}$ and $\bm{w}^* = \arginf_{\bm{w}}\norm{\wt{B}(\bm{w})}_{\nu}^2$. Define $\upsilon(\bm{w}^*) \triangleq \mathbb{E}_{\mathcal{N}(\bm{w}^*,\bm{I})}\left[ v(\bm{w}) \right]$, with $v(\bm{w}) \triangleq \mathbb{E}_{\nu}\left[Var_{\mathcal{P}}\left[b(\bm{w})\right]\right]$. Then, there exist constants $c_1,c_2,c_3,c_4$ such that, with probability at least $1-2\delta$ over the choice of weights $\bm{w} \sim \frac{1}{C}\sum_i\mathcal{N}(\wh{\bm{\mu}}_i,\wh{\bm{\Sigma}}_i)$ and dataset $D$:
\begin{equation}
\resizebox{0.93\textwidth}{!}{$
\norm{Q_{\bm{w}} - \wt{Q}}_{\nu}^2 \leq \frac{1}{(1-\gamma)\delta}\left(2\norm{\wt{B}(\bm{w}^{*})}_{\nu}^2 + \frac{c_1}{N^2} + \frac{c_2 + \frac{c_4}{\abs{\mathcal{W}_s}}\sum_{j}\norm{\bm{w}^* - \bm{w}_{j}}}{N} + \frac{\upsilon(\bm{w}^*) + c_3\sqrt{\log\frac{2}{\delta}}}{\sqrt{N}}\right)$}
\end{equation}
\end{restatable}
Similarly to the previous proof, we can apply Lemma \ref{lemma:general-bound} with variational parameters $\wh{\bm{\xi}} = (\wh{\bm{\mu}}_1,\dots,\wh{\bm{\mu}}_C,\wh{\bm{\Sigma}}_1,\dots,\wh{\bm{\Sigma}}_C)$, while choosing the same specific parameters for the right-hand side: $\bm{\mu}_i = \bm{w}^*$ and $\bm{\Sigma}_i = c\bm{I}$ for all $i=1,\dots,C$. Then, we obtain:
\begin{align}\label{eq:thmg-1}
\mathbb{E}_{q_{\wh{\bm{\xi}}}}\left[ \norm{B(\bm{w})}_{\nu}^2 \right ] \leq\ & \inf_{\bm{\xi} \in \Xi}\left\{ \mathbb{E}_{q_{\bm{\xi}}}\left[ \norm{B(\bm{w})}_{\nu}^2 \right ] + \mathbb{E}_{q_{\bm{\xi}}}\left[ v(\bm{w}) \right] + 2\frac{\lambda}{N} KL({q_{\bm{\xi}}}||p) \right\} + 2\frac{R_{max}^2}{(1-\gamma)^2}\sqrt{\frac{\log\frac{2}{\delta}}{N}}\notag\\ \leq\ & \mathbb{E}_{\mathcal{N}(\bm{w}^*,c\bm{I})}\left[ \norm{B(\bm{w})}_{\nu}^2 \right ] + \mathbb{E}_{\mathcal{N}(\bm{w}^*,c\bm{I})}\left[ v(\bm{w}) \right] + 2\frac{\lambda}{N} KL\left(\mathcal{N}(\bm{w}^*,c\bm{I})\ ||\ p\right)\notag\\ &\ \ + 2\frac{R_{max}^2}{(1-\gamma)^2}\sqrt{\frac{\log\frac{2}{\delta}}{N}}
The only difference w.r.t. Eq. \eqref{eq:thg-4} of Thm. \ref{th:main-gaussian} is the KL divergence term, which now contains a mixture distribution. From Thm. \ref{th:ukl} we have:
\begin{align}\label{eq:thmg-2}
KL(\mathcal{N}(\bm{w}^*,c\bm{I})\ ||\ p) \leq KL(\chi^{(2)}||\chi^{(1)}) + \sum_{j} \chi^{(2)}_{j}KL(\mathcal{N}(\bm{w}^*,c\bm{I})\ ||\ \mathcal{N}(\bm{w}_j, \sigma_p^2\bm{I}))
where the vectors $\chi^{(1)}$ and $\chi^{(2)}$ are the ones defined in Thm. \ref{th:ukl}. Notice that, since we reduced the posterior to one component, the can get rid of the index $i$. Using the definitions of these two vectors from Sec. 8 of \cite{hershey2007approximating}, we have:
\chi^{(1)}_j = \frac{1}{\abs{\mathcal{W}_s}}\ \forall j = 1,\dots,\abs{\mathcal{W}_s}
\begin{align}\label{eq:chi2}
\chi^{(2)}_j = \frac{e^{-KL(\mathcal{N}(\bm{w}^*,c\bm{I})\ ||\ \mathcal{N}(\bm{w}_j, \sigma_p^2\bm{I}))}}{\sum_{j'}e^{-KL(\mathcal{N}(\bm{w}^*,c\bm{I})\ ||\ \mathcal{N}(\bm{w}_{j'}, \sigma_p^2\bm{I}))}}\ \forall j = 1,\dots,\abs{\mathcal{W}_s}
Since the KL divergence is:
KL(\mathcal{N}(\bm{w}^*,c\bm{I})\ ||\ \mathcal{N}(\bm{w}_j, \sigma_p^2\bm{I})) = \frac{1}{2}\left( d\log\frac{\sigma^2_p}{c} + d\frac{c}{\sigma^2_p} + \frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_j} - d\right)
Eq. \ref{eq:chi2} can be rewritten as:
\chi^{(2)}_j = \frac{e^{-\frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_j}}}{\sum_{j'}e^{-\frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_{j'}}}}\ \forall j = 1,\dots,\abs{\mathcal{W}_s}
Let us bound the two terms of \eqref{eq:thmg-2} separately. For the first one, we have:
KL(\chi^{(2)}||\chi^{(1)}) =\ & \sum_j \chi^{(2)}_j \log\frac{\chi^{(2)}_j}{\chi^{(1)}_j}\\ =\ & \sum_j \chi^{(2)}_j \log\chi^{(2)}_j - \sum_j \chi^{(2)}_j \log\frac{1}{\abs{\mathcal{W}_s}}\\ =\ & \sum_j \chi^{(2)}_j \log\frac{e^{-\frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_j}}}{\sum_{j'}e^{-\frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_{j'}}}} - \log\frac{1}{\abs{\mathcal{W}_s}}\\ =\ & -\sum_j \chi^{(2)}_j \frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_j} - \sum_j \chi^{(2)}_j \log\sum_{j'}e^{-\frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_{j'}}} - \log\frac{1}{\abs{\mathcal{W}_s}}\\ =\ & -\sum_j \chi^{(2)}_j \frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_j} - \log\frac{1}{\abs{\mathcal{W}_s}}\sum_{j'}e^{-\frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_{j'}}}\\ \leq\ & -\sum_j \chi^{(2)}_j \frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_j} + \frac{1}{\abs{\mathcal{W}_s}}\sum_{j'}\frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_{j'}}
where we applied Jensen's inequality in the last step ($-\log$ is a convex function). The second term of \eqref{eq:thmg-2} is:
\sum_{j} \chi^{(2)}_{j}KL(\mathcal{N}(\bm{w}^*,c\bm{I})\ ||\ \mathcal{N}(\bm{w}_j, \sigma_p^2\bm{I})) =\ & \frac{1}{2}\sum_{j} \chi^{(2)}_{j}\left( d\log\frac{\sigma^2_p}{c} + d\frac{c}{\sigma^2_p} + \frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_j} - d\right)\\ \leq\ & \frac{1}{2}d\log\frac{\sigma^2_p}{c} + \frac{1}{2}d\frac{c}{\sigma^2_p} + \frac{1}{2}\sum_{j} \chi^{(2)}_{j}\frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_j}
Putting it all together:
KL(\mathcal{N}(\bm{w}^*,c\bm{I})\ ||\ p) \leq\ & \frac{1}{\abs{\mathcal{W}_s}}\sum_{j}\frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_{j}} -\frac{1}{2}\sum_j \chi^{(2)}_j \frac{1}{\sigma^2_p}\norm{\bm{w}^* - \bm{w}_j} + \frac{1}{2}d\log\frac{\sigma^2_p}{c}  + \frac{1}{2}d\frac{c}{\sigma^2_p}\\ \leq\ & \frac{1}{\sigma^2_p\abs{\mathcal{W}_s}}\sum_{j}\norm{\bm{w}^* - \bm{w}_{j}} + \frac{1}{2}d\log\frac{\sigma^2_p}{c}  + \frac{1}{2}d\frac{c}{\sigma^2_p}
Notice that, from now on, one can simply apply the proof of Thm. \ref{th:main-gaussian} with $\sigma_{max}=\sigma_{min}=\sigma^2_p$ and $\frac{1}{2}\norm{\bm{w}^* - \bm{\mu}_p}_{\bm{\Sigma}_p^{-1}} = \frac{1}{\sigma^2_p\abs{\mathcal{W}_s}}\sum_{j}\norm{\bm{w}^* - \bm{w}_{j}}$. Thus, by changing the three constants to $c_1 = \frac{1}{2}\gamma^2\kappa^2\bm{\phi}_{max}^2 + \frac{\lambda d}{\sigma^2_p}$, $c_2 = \bm{w}_{max}^2\bm{\phi}_{max}^2(1 + \gamma)^2 + \lambda d \log(\sigma^2_p + \log N)$\todo{$c_2$ actually contains $N$. Should we make this more explicit?}, $c_3 = 2\frac{R_{max}^2}{(1-\gamma)^2}$, and setting $c_4 = \frac{2\lambda}{\sigma^2_p}$, we can write that, with probability at least $1-2\delta$:
\norm{Q_{\bm{w}} - \wt{Q}}_{\nu}^2 \leq \frac{1}{(1-\gamma)\delta}\left(2\norm{\wt{B}(\bm{w}^{*})}_{\nu}^2 + \frac{c_1}{N^2} + \frac{c_2 + \frac{c_4}{\abs{\mathcal{W}_s}}\sum_{j}\norm{\bm{w}^* - \bm{w}_{j}}}{N} + \frac{\upsilon(\bm{w}^*) + c_3\sqrt{\log\frac{2}{\delta}}}{\sqrt{N}}\right)
